Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-21 NEUROC/sec2-review-of-graph-neural-network.tex
更新第二章

By default, GNNs are trained in a full-batch way, using the whole graph in each iteration.
The full-batch gradient descent has two disadvantages \cite{chiang2019_cluster_gcn}.
It has to cache intermediate results of all vertices in the forward phase, which consumes lots of memory space.
It updates the parameters only once for each epoch, slowing the convergence of gradient descent.

To train GNNs in a mini-batch way, the sampling techniques are proposed.
In each mini-batch, they sample a small subgraph from the whole graph $\mathcal{G}$ and uses the subgraph to update the model parameters.
The sampling techniques only active the graph neurons and the connections that appear in the sampled subgraph between GNN layers, as shown in \figurename~\ref{fig:gnn_sampling}.
In active graph neurons and connections do not participate in the training of this mini-batch, saving lots of computation and storage costs.
Moreover, it may reduce the risk of overfitting the training graph.
The existing sampling techniques can be classified into two groups \textbf{neighbor sampling and graph sampling} based on whether different GNN layers sample different subgraphs.

The neighbor sampling techniques \cite{hamilton2017_graphsage, ying2018_pinsage, chen2018_fastgcn, chen2018_sgcn, huang2018_adap} sample nodes or edges layer by layer.
The sampled subgraphs of different GNN layers may be \emph{different}, as shown in \figurename~\ref{fig:gnn_sampling_neighbor_sampling}.
GraphSAGE \cite{hamilton2017_graphsage} is the representative technique.
The tecnnique first samples several vertices from $\mathcal{V}$ in the last GNN layer.
Then it repeatedly samples the neighbors of those sampled vertices in the previous layer until the input layer.
For every sampled vertex $v_x$ in the GNN layer $l$, GraphSAGE samples at most $S^l$ neighbors of $v_x$ from the previous GNN layer.
$S^l$ is the hyper-parameter that is usually much smaller than $|\mathcal{V}|$.
In this way, GraphSAGE limits the neighborhood sizes of the vertices in the sampled subgraph, especially high-degree vertices.

\begin{figure}
    \centering
    \subfloat[Neighbor sampling\label{fig:gnn_sampling_neighbor_sampling}]{\includegraphics[height=4.5cm]{figs/illustration/layer_sampling.pdf}}
    %
    \subfloat[Graph sampling\label{fig:gnn_sampling_graph_sampling}]{\includegraphics[height=4.5cm]{figs/illustration/graph_sampling.pdf}}
    \caption{Training a GNN with sampling techniques. The faded graph neurons and their connections are inactivated.}
    \label{fig:gnn_sampling}
\end{figure}

The graph sampling techniques \cite{zeng2018_aesg, chiang2019_cluster_gcn, zeng2020_graphsaint} sample a subgraph for each mini-batch and use the same sampled graph for all GNN layers, as shown in \figurename~\ref{fig:gnn_sampling_graph_sampling}.
They differ in the methods to sample subgraphs.
The cluster sampler technique \cite{chiang2019_cluster_gcn} is the representative technique.
Given a training graph $\mathcal{G}$, it partitions $\mathcal{G}$ into dense clusters.
For each mini-batch, it randomly pick $N$ clusters to form the sampled subgraph, where $N$ is the hyper-parameter.

>>>>

By default, GNNs are trained in a full-batch way, using the whole graph in each iteration.
%
The full-batch gradient descent has two disadvantages \cite{chiang2019_cluster_gcn}:
%
(1) it has to cache intermediate results of all vertices in the forward phase, which consumes lots of memory space;
%
(2) it updates the model parameters only once for each epoch, slowing the convergence of gradient descent.

To train GNNs in a mini-batch way, the sampling techniques \cite{hamilton2017_graphsage, ying2018_pinsage, chen2018_fastgcn, chen2018_sgcn, zeng2018_aesg, chiang2019_cluster_gcn, zeng2020_graphsaint} are proposed.
%
In each mini-batch, they sample a small subgraph from the whole graph $\mathcal{G}$ and uses the subgraph to update the model parameters.
%
The sampling techniques only active the graph neurons and the connections that appear in the sampled subgraph between GNN layers, as shown in \figurename~\ref{fig:gnn_sampling}.
%
Inactive graph neurons and connections do not participate in the training of this mini-batch, saving lots of computation and storage costs.
%
Moreover, it may reduce the risk of overfitting the training graph.
%
The existing sampling techniques can be classified into two groups: \emph{neighbor sampling} and \emph{graph sampling}, based on whether different GNN layers sample different subgraphs.

The neighbor sampling techniques \cite{hamilton2017_graphsage, ying2018_pinsage, chen2018_fastgcn, chen2018_sgcn, huang2018_adap} sample nodes or edges layer by layer.
%
The sampled subgraphs of different GNN layers may be \emph{different}, as shown in \figurename~\ref{fig:gnn_sampling}(a).
%
GraphSAGE \cite{hamilton2017_graphsage} is a representative neighbor sampling technique.
%
The tecnnique first samples several vertices from $\mathcal{V}$ in the last GNN layer.
%
Then it repeatedly samples the neighbors of the sampled vertices from the previous layer until the input layer.
%
For every sampled vertex $v_x$ in the GNN layer $l$, GraphSAGE samples at most $\eta^l$ neighbors of $v_x$ from the previous layer.
%
$\eta^l$ is the hyper-parameter that is usually much smaller than $|\mathcal{V}|$.
%
By this way, GraphSAGE limits neighborhood sizes of vertices in the sampled subgraph, especially high-degree vertices.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\columnwidth]{figs/illustration/sampling_techniques.pdf}
    %\subfloat[Neighbor sampling\label{fig:gnn_sampling_neighbor_sampling}]{\includegraphics[height=4.5cm]{figs/illustration/layer_sampling.pdf}}
    %
    %\subfloat[Graph sampling\label{fig:gnn_sampling_graph_sampling}]{\includegraphics[height=4.5cm]{figs/illustration/graph_sampling.pdf}}
    \caption{Training a GNN with sampling techniques. The faded graph neurons and their connections are inactivated.}
    \label{fig:gnn_sampling}
\end{figure}

The graph sampling techniques \cite{zeng2018_aesg, chiang2019_cluster_gcn, zeng2020_graphsaint} sample a subgraph for each mini-batch and use the same sampled subgraph for all GNN layers, as shown in \figurename~\ref{fig:gnn_sampling}(b).
%
They differ in how to sample subgraphs.
%
The cluster sampler technique \cite{chiang2019_cluster_gcn} is a representative graph sampling technique.
%
Given a training graph $\mathcal{G}$, it partitions $\mathcal{G}$ into several dense clusters.
%
For each mini-batch, it randomly pick $N$ clusters to form the sampled subgraph, where $N$ is the hyper-parameter.

### 2020-12-21 NEUROC/sec4-experiments.tex
更新第二章

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_pubmed.pdf}
    \includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-photo.pdf}
    \includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_coauthor-physics.pdf}

    \includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-computers.pdf}
    \includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_flickr.pdf}
    \includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_com-amazon.pdf}
    %\subfloat[\label{fig:exp_absolute_training_time_pubmed}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_pubmed.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_amazon-photo}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-photo.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_coauthor-physics}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_coauthor-physics.pdf}} \\
    %\subfloat[\label{fig:exp_absolute_training_time_amazon-computers}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-computers.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_flickr}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_flickr.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_com-amazon}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_com-amazon.pdf}}
    \caption{Distribution of the wall-clock training time of 50 epoches on different datasets. GaAN crashed due to out of memory exception on the \texttt{cph} dataset.}
    \label{fig:exp_absolute_training_time}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gcn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GCN.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_ggnn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GGNN.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gat.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GAT.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat}
\end{figure}

>>>>

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.3\textwidth]{figs/experiments/exp_absolute_training_time_comparison_pubmed.pdf}
    \includegraphics[width=0.3\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-photo.pdf}
    \includegraphics[width=0.3\textwidth]{figs/experiments/exp_absolute_training_time_comparison_coauthor-physics.pdf}

    \includegraphics[width=0.3\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-computers.pdf}
    \includegraphics[width=0.3\textwidth]{figs/experiments/exp_absolute_training_time_comparison_flickr.pdf}
    \includegraphics[width=0.3\textwidth]{figs/experiments/exp_absolute_training_time_comparison_com-amazon.pdf}
    %\subfloat[\label{fig:exp_absolute_training_time_pubmed}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_pubmed.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_amazon-photo}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-photo.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_coauthor-physics}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_coauthor-physics.pdf}} \\
    %\subfloat[\label{fig:exp_absolute_training_time_amazon-computers}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_amazon-computers.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_flickr}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_flickr.pdf}}
    %\subfloat[\label{fig:exp_absolute_training_time_com-amazon}]{\includegraphics[width=0.32\textwidth]{figs/experiments/exp_absolute_training_time_comparison_com-amazon.pdf}}
    \caption{Distribution of the wall-clock training time of 50 epoches on different datasets. GaAN crashed due to out of memory exception on the \texttt{cph} dataset.}
    \label{fig:exp_absolute_training_time}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gcn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GCN.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_ggnn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GGNN.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gat.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GAT.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat}
\end{figure}
